---
title: "CSLS Water Chem Visualization"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Put the title of your vignette here}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, echo=FALSE, warning=FALSE, message=FALSE}
library(knitr)
library(CSLSchem)
library(dplyr)
library(stringr)
library(lubridate)
library(reshape2)
library(extrafont)

text_size    <- 12
lakes        <- c("Pleasant", "Long", "Plainfield")
SWIMS        <- CSLSchem::SWIMS

find_and_plot <- function(parameter, plotting_name, text_size) {
  SWIMS_parameter <- filter_parameter(parameter, plotting_name)
  plot_obj        <- plot_overall(SWIMS_parameter, text_size)
  return(plot_obj)
}

plot_timeseries_fcn <- function(parameter, plotting_name, text_size, 
                                static_gw = TRUE) {
  SWIMS_parameter <- filter_parameter(parameter = parameter, 
                                      plotting_name = plotting_name,
                                      static_gw = static_gw)
  plot_obj <- plot_timeseries(SWIMS_parameter, text_size)
  return(plot_obj)
}

```

# Overview {#top}
The data used here includes data from both SWIMS and LDES. Duplicates between
the two data sources as well as samples labeled as blanks or duplicates are
removed. Overall notes:

* Seems probable that Long Lake and Plainfield Lake sample labels or data entry
at WSLH were swapped on July 16, 2019 (see calcium, magnesium, alkalinity,
conductivity, total recoverable hardness).


## Calcium, Total Recoverable
[Back to top](#top)

Observations:

* Calcium is fairly stable over time at lake and groundwater sites except for
one wonky well (PFL-09).
* Lakes are generally lower than groundwater.
* Pleasant and Plainfield lakes are similar, Long lake is notably lower.
* Upgradient and downgradient groundwater behave similarly at Long and
Plainfield, downgradient groundwater is a little more similar to lake water at
Pleasant.
  
```{r Ca, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=5, fig.align="center"}
parameter     <-"CALCIUM TOTAL RECOVERABLE"
plotting_name <- "Total Recoverable Calcium"

SWIMS_parameter <- filter_parameter(SWIMS, parameter, plotting_name)
plot_summary(SWIMS_parameter, text_size, yaxis_type = "log")
plot_timeseries(SWIMS_parameter, text_size, yaxis_type = "log")
```

## Magnesium, Total Recoverable
[Back to top](#top)

Observations:

* Magnesium is fairly stable over time at lake and groundwater sites except for
one wonky well (PFL-09).
* Lakes are generally lower than groundwater.
* Pleasant and Plainfield lakes are similar, Long lake is notably lower.
* Upgradient and downgradient groundwater behave similarly at Long and
Plainfield, downgradient groundwater is a little more similar to lake water at
Pleasant.

```{r Mg, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=5, fig.align="center"}
parameter     <- "MAGNESIUM TOTAL RECOVERABLE"
plotting_name <- "Total Recoverabel Magnesium"

SWIMS_parameter <- filter_parameter(SWIMS, parameter, plotting_name)
plot_summary(SWIMS_parameter, text_size, yaxis_type = "log")
plot_timeseries(SWIMS_parameter, text_size, yaxis_type = "log")
```

## Alkalinity, Total CaCO3
[Back to top](#top)

Observations:

* Alkalinity is fairly stable over time at lake and groundwater sites except for
one wonky well (PFL-09).
* Lakes are generally lower than groundwater.
* Pleasant and Plainfield lakes are similar, Long lake is notably lower.
* Upgradient and downgradient groundwater behave similarly at Long and
Plainfield, downgradient groundwater is a little more similar to lake water at
Pleasant.

```{r CACO3, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=5, fig.align="center"}
parameter     <-"ALKALINITY TOTAL CACO3"
plotting_name <- "Alkalinity Total CaCO3"

SWIMS_parameter <- filter_parameter(SWIMS, parameter, plotting_name)
plot_summary(SWIMS_parameter, text_size, yaxis_type = "log")
plot_timeseries(SWIMS_parameter, text_size, yaxis_type = "log")
```


## Hardness, Total Recoverable Calculation
[Back to top](#top)

Observations:

* Hardness is fairly stable over time at lake and groundwater sites except for
one wonky well (PFL-09).
* Lakes are generally lower than groundwater.
* Pleasant and Plainfield lakes are similar, Long lake is notably lower.
* Upgradient and downgradient groundwater behave similarly at Long and
Plainfield, downgradient groundwater is a little more similar to lake water at
Pleasant.

```{r hardness, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=5, fig.align="center"}
parameter     <-"HARDNESS TOTAL RECOVERABLE CALCULATION"
plotting_name <- "Total Recoverable Hardness"

SWIMS_parameter <- filter_parameter(SWIMS, parameter, plotting_name)
plot_summary(SWIMS_parameter, text_size, yaxis_type = "log")
plot_timeseries(SWIMS_parameter, text_size, yaxis_type = "log")
```

## Aluminum, Total Recoverable
[Back to top](#top)

Observations:

* Aluminum is s variable within groundwater sites as it is among groundwater sites.
* Lake levels are very low compared to groundwater levels (typically below
detection).
* At Pleasant Lake, downgradient wells show a downward trend in aluminum 
  over time.
* At Long Lake and Plainfield Lake, there is no discernable trend and there is 
  as much variation within well samples as there is among well samples.

```{r Al, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=5, fig.align="center"}
parameter     <-"ALUMINUM,TOTAL RECOVERABLE"
plotting_name <- "Total Recoverable Aluminum"

SWIMS_parameter <- filter_parameter(SWIMS, parameter, plotting_name)
plot_summary(SWIMS_parameter, text_size, yaxis_type = "log")
plot_timeseries(SWIMS_parameter, text_size, yaxis_type = "log")
```

## Iron, Total Recoverable
[Back to top](#top)

Observations:

* Iron is as variable within groundwater sites as it is among groundwater sites.
* Groundwater levels are generally low (near or below the limit of quality) with
the exception of one well at Pleasant (PSNT-05) and one well at Plainfield
(PFL-09).
* Iron was only measured at the lake surfaces one time (in Feb 2019) and was
below detection at Pleasant, below the limit of quality at Plainfield, and less
than 1 mg/L at Long.

```{r Fe, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=5, fig.align="center"}
parameter     <-"IRON TOTAL RECOVERABLE"
plotting_name <- "Total Recoverable Iron"

SWIMS_parameter <- filter_parameter(SWIMS, parameter, plotting_name)
plot_summary(SWIMS_parameter, text_size, yaxis_type = "log")
plot_timeseries(SWIMS_parameter, text_size, yaxis_type = "log")
```

## Manganese, Total Recoverable
[Back to top](#top)

Observations:

* Manganese is as variable within groundwater sites as it is among groundwater
sites.
* Groundwater levels are similar at all three lakes with the exception of one
well at Pleasant (PSNT-05) and one well at Plainfield (PFL-09) which are notably
higher.
* While there are very few lake samples, they appear to be similar to
groundwater concentrations.

```{r Mn, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=5, fig.align="center"}
parameter     <-"MANGANESE, TOTAL RECOVERABLE"
plotting_name <- "Total Recoverable Manganese"

SWIMS_parameter <- filter_parameter(SWIMS, parameter, plotting_name)
plot_summary(SWIMS_parameter, text_size, yaxis_type = "log")
plot_timeseries(SWIMS_parameter, text_size, yaxis_type = "log")
```

## Conductivity
[Back to top](#top)

Observations:

* Conductivity is fairly stable over time at lake and groundwater sites.
* Lakes are generally lower than groundwater
* Conductivity is similar at Pleasant Lake and Plainfield Lake, but
substantially lower at Long Lake.
* Upgradient and downgradient groundwater behave similarly, though downgradient
groundwater is a little closer to lake water at Pleasant than at the other
lakes.


```{r Conductivity, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=5, fig.align="center"}
parameter     <-"CONDUCTIVITY, UMHOS/CM @ 25C"
plotting_name <- "Conductivity"

SWIMS_parameter <- filter_parameter(SWIMS, parameter, plotting_name)
plot_summary(SWIMS_parameter, text_size, yaxis_type = "log")
plot_timeseries(SWIMS_parameter, text_size, yaxis_type = "log")
```

## Chloride
[Back to top](#top)

Observations:

* Lakes are less variable than groundwater
* Groundwater is about as variable within as site as across sites
* Lake values and groundwater values are similar at Pleasant and Long, but
Plainfield lake values are notably higher than Plainfield groundwater.
* Lake and gorundwater values are lowest at Long - nearly all samples are below
the limit of quality.
* Do we have a measure of road density or impervious surfaces around each lake
(say a 200m buffer)?
* I think all concentrations are well below levels of concern (that would be 200
mg/l, correct)?

```{r Cl, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=5, fig.align="center"}
parameter     <-"CHLORIDE"
plotting_name <- "Chloride"

SWIMS_parameter <- filter_parameter(SWIMS, parameter, plotting_name)
plot_summary(SWIMS_parameter, text_size, yaxis_type = "log")
plot_timeseries(SWIMS_parameter, text_size, yaxis_type = "log")
```

## Sodium, Total Recoverable
[Back to top](#top)

Observations:

* Sodium is fairly stable over time at lake and groundwater sites
* Lake values are generally lower than groundwater values at Pleasant and Long,
but generally higher than groundwater at Plainfield.
* Groundwater and especially lake values are lowest at Long Lake.

```{r Na, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=5, fig.align="center"}
parameter     <-"SODIUM TOTAL RECOVERABLE"
plotting_name <- "Total Recoverable Sodium"

SWIMS_parameter <- filter_parameter(SWIMS, parameter, plotting_name)
plot_summary(SWIMS_parameter, text_size, yaxis_type = "log")
plot_timeseries(SWIMS_parameter, text_size, yaxis_type = "log")
```

## Nitrogen, NH3-N Dissolved
[Back to top](#top)

Observations:

* NH3-N is behaves similarly at all three lakes
* NH3-N is almost non-existant in groundwater (at/below the limit of quality in
all but two samples) and still very low in lakes

```{r NH3, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=5, fig.align="center"}
parameter     <-"NITROGEN NH3-N DISS"
plotting_name <- "Dissolved Ammonium as Nitrogen"

SWIMS_parameter <- filter_parameter(SWIMS, parameter, plotting_name)
plot_summary(SWIMS_parameter, text_size)
plot_timeseries(SWIMS_parameter, text_size)
```

## Nitrogen, NO3 + NO2 Dissolved (as N)
[Back to top](#top)

Observations:

* All sites have at least one well with appreciable NO3/NO2, but most
groundwater and lake samples are near the limit of quality.
* Pleasant Lake has slightly higher (i.e., usually measureable) lake values
than Long and Plainfield, which are all non-detect.

```{r NO3NO2, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=5, fig.align="center"}
parameter     <-"NITROGEN NO3+NO2 DISS (AS N)"
plotting_name <- "Dissolved Nitrate/Nitrite as Nitrogen"

SWIMS_parameter <- filter_parameter(SWIMS, parameter, plotting_name)
plot_summary(SWIMS_parameter, text_size)
plot_timeseries(SWIMS_parameter, text_size)
```

## Phosphorus, Total
[Back to top](#top)

Observations:

* Phoshorous is anomalously high at one Pleasant well (PSNT-05) and one
Plainfield well (PFL-09).
* Otherwise, groundwater levels are typically similar to or slightly higher than
lake levels
* Lake values are often below the limit of quality, particularly at Pleasant
Lake.

```{r P, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=5, fig.align="center"}
parameter     <-"PHOSPHORUS TOTAL"
plotting_name <- "Total Phosphorus"

SWIMS_parameter <- filter_parameter(SWIMS, parameter, plotting_name)
plot_summary(SWIMS_parameter, text_size, yaxis_type = "log")
plot_timeseries(SWIMS_parameter, text_size, yaxis_type = "log")
```

## Potassium, Total Recoverable
[Back to top](#top)

Observations:

* Lake values are typically lower than groundwater levels at Pleasant, but
similar to groundwater levels at Long and Plainfield
* Plainfield has several wells with notably higher levels, including PFL-09.

```{r K, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=5, fig.align="center"}
parameter     <-"POTASSIUM TOTAL RECOVERABLE"
plotting_name <- "Total Recoverable Potassium"

SWIMS_parameter <- filter_parameter(SWIMS, parameter, plotting_name)
plot_summary(SWIMS_parameter, text_size, yaxis_type = "log")
plot_timeseries(SWIMS_parameter, text_size, yaxis_type = "log")
```

## Silica, Dissolved (as SiO2)
[Back to top](#top)

Observations:

* Lake values may have a seasonal pattern
* Lake vlaues are lower than groundwater values
* Downgradient values are lower than upgradinet values at Pleasant Lake, but all
groundwater is similar at Long and Plainfield.
* Levels are less variable in grundwater than in the lake

```{r SiO2, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=5, fig.align="center"}
parameter     <-"SILICA, DISSOLVED (MG/L AS SI02)"
plotting_name <- "Dissolved Silica as SiO2"

SWIMS_parameter <- filter_parameter(SWIMS, parameter, plotting_name)
plot_summary(SWIMS_parameter, text_size)
plot_timeseries(SWIMS_parameter, text_size)
```

## pH
[Back to top](#top)

Observations:

* Lakes and groundwater are about as variable among sites as within sites
* Lake pH is higher than groundwater pH at Pleasant and Plainfield, but
lower at Long.
* There is an anomolously low value for LL-05 from the last sample date. 

```{r pH, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=5, fig.align="center"}
parameter     <-"PH LAB"
plotting_name <- "pH"

SWIMS_parameter <- filter_parameter(SWIMS, parameter, plotting_name)
plot_summary(SWIMS_parameter, text_size)
plot_timeseries(SWIMS_parameter, text_size)
```


## Sulfate, Total
[Back to top](#top)

Observations:

* Sulfate is below the limit of quality for nearly all samples

```{r So4, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=5, fig.align="center"}
parameter     <-"SULFATE TOTAL"
plotting_name <- "Total Sulfate"

SWIMS_parameter <- filter_parameter(SWIMS, parameter, plotting_name)
plot_summary(SWIMS_parameter, text_size)
plot_timeseries(SWIMS_parameter, text_size)
```

