---
title: "CSLS Water Chem Visualization"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Put the title of your vignette here}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, echo=FALSE, warning=FALSE, message=FALSE}
library(knitr)
library(CSLSchem)
library(dplyr)
library(stringr)
library(lubridate)
library(reshape2)
library(extrafont)

text_size   <- 12
lakes       <- c("Pleasant", "Long", "Plainfield")
water_chem  <- CSLSdata::water_chem
water_chem  <- water_chem %>%
               filter(!site_type %in% c("precipitation", "deep"))
site_types  <- c("upgradient", "nogradient", "downgradient", "lake", "lake_bottom")
site_labels <- c("Upgr.", "Nogr.", "Downgr.", "Lake Surf", "Lake Bot.")
site_colors <- c("#33A02C", "#B15928", "#B2DF8A", "#FB9A99", "#E31A1C")

```

# Overview {#top}
The data used here includes data from both SWIMS and LDES. Duplicates between
the two data sources as well as samples labeled as blanks or duplicates are
removed.

## Calcium, Total Recoverable
[Back to top](#top)

Observations:

* Calcium is fairly stable over time at lake and groundwater sites except for
one wonky well (PFL-09).
* Lakes are generally lower than groundwater.
* Pleasant and Plainfield lakes are similar, Long lake is notably lower.
* Upgradient and downgradient groundwater behave similarly at Long and
Plainfield, downgradient groundwater is a little more similar to lake water at
Pleasant.
  
```{r Ca, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=5, fig.align="center"}
parameter     <-"CALCIUM TOTAL RECOVERABLE"
plotting_name <- "Total Recoverable Calcium"

this_parameter <- filter_parameter(water_chem, parameter, plotting_name)
plot_summary(this_parameter, yaxis_type = "log", text_size, 
             site_types, site_labels, site_colors)
plot_timeseries(this_parameter, yaxis_type = "log", text_size, 
                site_types, site_labels, site_colors)
```

## Magnesium, Total Recoverable
[Back to top](#top)

Observations:

* Magnesium is fairly stable over time at lake and groundwater sites except for
one wonky well (PFL-09).
* Lakes are generally lower than groundwater.
* Pleasant and Plainfield lakes are similar, Long lake is notably lower.
* Upgradient and downgradient groundwater behave similarly at Long and
Plainfield, downgradient groundwater is a little more similar to lake water at
Pleasant.

```{r Mg, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=5, fig.align="center"}
parameter     <- "MAGNESIUM TOTAL RECOVERABLE"
plotting_name <- "Total Recoverabel Magnesium"

this_parameter <- filter_parameter(water_chem, parameter, plotting_name)
plot_summary(this_parameter, yaxis_type = "log", text_size, 
             site_types, site_labels, site_colors)
plot_timeseries(this_parameter, yaxis_type = "log", text_size, 
                site_types, site_labels, site_colors)
```

## Ca:Mg Ratio
[Back to top](#top)

Observations:

* At all lakes, there is dip in Ca:Mg values from spring to summer. This happens
a bit earlier in Long Lake and Plainfield Lake (~March-June) than it does at
Pleasant Lake (~April-August).
* Most wells also have a small dip in Ca:Mg relative to their overall timeseries
in June 2019, particularly at Long Lake.
* At Pleasant Lake, Ca:MG is similar at lake and groundwater sites. At Long
Lake, the lake is at the high end of the range in groundwater sites. At
Plainfield Lake, the lake is generally higher than groundwater sites.
* There is no pattern in whether upgradient or downgradient wells have a higher Ca:Mg
* Ca:Mg is more variable spatially than temporally in groundwater, and is
generally more stable in groundwater than at lakes.

```{r Ca_Mg, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=5, fig.align="center"}
parameter     <-"CALCIUM TOTAL RECOVERABLE"
plotting_name <- "Total Recoverable Calcium"
Ca            <- filter_parameter(water_chem, parameter, plotting_name)
Ca            <- Ca %>%
                 select(.data$lake, .data$site_type, .data$date, .data$result,
                        .data$site_id)

parameter     <- "MAGNESIUM TOTAL RECOVERABLE"
plotting_name <- "Total Recoverabel Magnesium"
Mg            <- filter_parameter(water_chem, parameter, plotting_name)
Mg            <- Mg %>%
                 select(.data$lake, .data$site_type, .data$date, .data$result,
                        .data$site_id)

Ca_Mg         <- merge(Ca, Mg, by = c("date", "lake", "site_type", "site_id"))
Ca_Mg$result  <- Ca_Mg$result.x/Ca_Mg$result.y
Ca_Mg$lod     <- 0
Ca_Mg$loq     <- 0

plot_summary(Ca_Mg, yaxis_type = "number", text_size, 
             site_types, site_labels, site_colors)
plot_timeseries(Ca_Mg, yaxis_type = "number", text_size, 
                site_types, site_labels, site_colors)
```

## Alkalinity, Total CaCO3
[Back to top](#top)

Observations:

* Alkalinity is fairly stable over time at lake and groundwater sites except for
one wonky well (PFL-09).
* Lakes are generally lower than groundwater.
* Pleasant and Plainfield lakes are similar, Long lake is notably lower.
* Upgradient and downgradient groundwater behave similarly at Long and
Plainfield, downgradient groundwater is a little more similar to lake water at
Pleasant.

```{r CACO3, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=5, fig.align="center"}
parameter     <- c("ALKALINITY TOTAL CACO3", "ALKALINITY TOTAL GRAN AVAL UEQ/L")
plotting_name <- "Alkalinity Total CaCO3"

this_parameter <- filter_parameter(water_chem, parameter, plotting_name)
plot_summary(this_parameter, yaxis_type = "log", text_size, 
             site_types, site_labels, site_colors)
plot_timeseries(this_parameter, yaxis_type = "log", text_size, 
                site_types, site_labels, site_colors)
```


## Hardness, Total Recoverable Calculation
[Back to top](#top)

Observations:

* Hardness is fairly stable over time at lake and groundwater sites except for
one wonky well (PFL-09).
* Lakes are generally lower than groundwater.
* Pleasant and Plainfield lakes are similar, Long lake is notably lower.
* Upgradient and downgradient groundwater behave similarly at Long and
Plainfield, downgradient groundwater is a little more similar to lake water at
Pleasant.

```{r hardness, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=5, fig.align="center"}
parameter     <-"HARDNESS TOTAL RECOVERABLE CALCULATION"
plotting_name <- "Total Recoverable Hardness"

this_parameter <- filter_parameter(water_chem, parameter, plotting_name)
plot_summary(this_parameter, yaxis_type = "log", text_size, 
             site_types, site_labels, site_colors)
plot_timeseries(this_parameter, yaxis_type = "log", text_size, 
                site_types, site_labels, site_colors)
```

## Aluminum, Total Recoverable
[Back to top](#top)

Observations:

* Aluminum is s variable within groundwater sites as it is among groundwater sites.
* Lake levels are very low compared to groundwater levels (typically below
detection).
* At Pleasant Lake, downgradient wells show a downward trend in aluminum 
  over time.
* At Long Lake and Plainfield Lake, there is no discernable trend and there is 
  as much variation within well samples as there is among well samples.

```{r Al, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=5, fig.align="center"}
parameter     <-"ALUMINUM,TOTAL RECOVERABLE"
plotting_name <- "Total Recoverable Aluminum"

this_parameter <- filter_parameter(water_chem, parameter, plotting_name)
plot_summary(this_parameter, yaxis_type = "log", text_size, 
             site_types, site_labels, site_colors)
plot_timeseries(this_parameter, yaxis_type = "log", text_size, 
                site_types, site_labels, site_colors)
```

## Iron, Total Recoverable
[Back to top](#top)

Observations:

* Iron is as variable within groundwater sites as it is among groundwater sites.
* Groundwater levels are generally low (near or below the limit of quality) with
the exception of one well at Pleasant (PSNT-05) and one well at Plainfield
(PFL-09).
* Iron was only measured at the lake surfaces one time (in Feb 2019) and was
below detection at Pleasant, below the limit of quality at Plainfield, and less
than 1 mg/L at Long.

```{r Fe, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=5, fig.align="center"}
parameter     <-"IRON TOTAL RECOVERABLE"
plotting_name <- "Total Recoverable Iron"

this_parameter <- filter_parameter(water_chem, parameter, plotting_name)
plot_summary(this_parameter, yaxis_type = "log", text_size, 
             site_types, site_labels, site_colors)
plot_timeseries(this_parameter, yaxis_type = "log", text_size, 
                site_types, site_labels, site_colors)
```

## Manganese, Total Recoverable
[Back to top](#top)

Observations:

* Manganese is as variable within groundwater sites as it is among groundwater
sites.
* Groundwater levels are similar at all three lakes with the exception of one
well at Pleasant (PSNT-05) and one well at Plainfield (PFL-09) which are notably
higher.
* While there are very few lake samples, they appear to be similar to
groundwater concentrations.

```{r Mn, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=5, fig.align="center"}
parameter     <-"MANGANESE, TOTAL RECOVERABLE"
plotting_name <- "Total Recoverable Manganese"

this_parameter <- filter_parameter(water_chem, parameter, plotting_name)
plot_summary(this_parameter, yaxis_type = "log", text_size, 
             site_types, site_labels, site_colors)
plot_timeseries(this_parameter, yaxis_type = "log", text_size, 
                site_types, site_labels, site_colors)
```

## Conductivity
[Back to top](#top)

Observations:

* Conductivity is fairly stable over time at lake and groundwater sites.
* Lakes are generally lower than groundwater
* Conductivity is similar at Pleasant Lake and Plainfield Lake, but
substantially lower at Long Lake.
* Upgradient and downgradient groundwater behave similarly, though downgradient
groundwater is a little closer to lake water at Pleasant than at the other
lakes.


```{r Conductivity, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=5, fig.align="center"}
parameter     <-"CONDUCTIVITY, UMHOS/CM @ 25C"
plotting_name <- "Conductivity"

this_parameter <- filter_parameter(water_chem, parameter, plotting_name)
plot_summary(this_parameter, yaxis_type = "log", text_size, 
             site_types, site_labels, site_colors)
plot_timeseries(this_parameter, yaxis_type = "log", text_size, 
                site_types, site_labels, site_colors)
```

## Chloride
[Back to top](#top)

Observations:

* Lakes are less variable than groundwater
* Groundwater is about as variable within as site as across sites
* Lake values and groundwater values are similar at Pleasant and Long, but
Plainfield lake values are notably higher than Plainfield groundwater.
* Lake and gorundwater values are lowest at Long - nearly all samples are below
the limit of quality.
* Do we have a measure of road density or impervious surfaces around each lake
(say a 200m buffer)?
* I think all concentrations are well below levels of concern (that would be 200
mg/l, correct)?

```{r Cl, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=5, fig.align="center"}
parameter     <-"CHLORIDE"
plotting_name <- "Chloride"

this_parameter <- filter_parameter(water_chem, parameter, plotting_name)
plot_summary(this_parameter, yaxis_type = "log", text_size, 
             site_types, site_labels, site_colors)
plot_timeseries(this_parameter, yaxis_type = "log", text_size, 
                site_types, site_labels, site_colors)
```

## Sodium, Total Recoverable
[Back to top](#top)

Observations:

* Sodium is fairly stable over time at lake and groundwater sites
* Lake values are generally lower than groundwater values at Pleasant and Long,
but generally higher than groundwater at Plainfield.
* Groundwater and especially lake values are lowest at Long Lake.

```{r Na, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=5, fig.align="center"}
parameter     <-"SODIUM TOTAL RECOVERABLE"
plotting_name <- "Total Recoverable Sodium"

this_parameter <- filter_parameter(water_chem, parameter, plotting_name)
plot_summary(this_parameter, yaxis_type = "log", text_size, 
             site_types, site_labels, site_colors)
plot_timeseries(this_parameter, yaxis_type = "log", text_size, 
                site_types, site_labels, site_colors)
```

## Nitrogen, NH3-N Dissolved
[Back to top](#top)

Observations:

* NH3-N is behaves similarly at all three lakes
* NH3-N is almost non-existant in groundwater (at/below the limit of quality in
all but two samples) and still very low in lakes

```{r NH3, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=5, fig.align="center"}
parameter     <-"NITROGEN NH3-N DISS"
plotting_name <- "Dissolved Ammonium as Nitrogen"

this_parameter <- filter_parameter(water_chem, parameter, plotting_name)
plot_summary(this_parameter, yaxis_type = "number", text_size, 
             site_types, site_labels, site_colors)
plot_timeseries(this_parameter, yaxis_type = "number", text_size, 
                site_types, site_labels, site_colors)
```

## Nitrogen, NO3 + NO2 Dissolved (as N)
[Back to top](#top)

Observations:

* All sites have at least one well with appreciable NO3/NO2, but most
groundwater and lake samples are near the limit of quality.
* Pleasant Lake has slightly higher (i.e., usually measureable) lake values
than Long and Plainfield, which are all non-detect.

```{r NO3NO2, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=5, fig.align="center"}
parameter     <-"NITROGEN NO3+NO2 DISS (AS N)"
plotting_name <- "Dissolved Nitrate/Nitrite as Nitrogen"

this_parameter <- filter_parameter(water_chem, parameter, plotting_name)
plot_summary(this_parameter, yaxis_type = "number", text_size, 
             site_types, site_labels, site_colors)
plot_timeseries(this_parameter, yaxis_type = "number", text_size, 
                site_types, site_labels, site_colors)
```

## Phosphorus, Total
[Back to top](#top)

Observations:

* Phoshorous is anomalously high at one Pleasant well (PSNT-05) and one
Plainfield well (PFL-09).
* Otherwise, groundwater levels are typically similar to or slightly higher than
lake levels
* Lake values are often below the limit of quality, particularly at Pleasant
Lake.

```{r P, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=5, fig.align="center"}
parameter     <-"PHOSPHORUS TOTAL"
plotting_name <- "Total Phosphorus"

this_parameter <- filter_parameter(water_chem, parameter, plotting_name)
plot_summary(this_parameter, yaxis_type = "log", text_size, 
             site_types, site_labels, site_colors)
plot_timeseries(this_parameter, yaxis_type = "log", text_size, 
                site_types, site_labels, site_colors)
```

## Potassium, Total Recoverable
[Back to top](#top)

Observations:

* Lake values are typically lower than groundwater levels at Pleasant, but
similar to groundwater levels at Long and Plainfield
* Plainfield has several wells with notably higher levels, including PFL-09.

```{r K, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=5, fig.align="center"}
parameter     <-"POTASSIUM TOTAL RECOVERABLE"
plotting_name <- "Total Recoverable Potassium"

this_parameter <- filter_parameter(water_chem, parameter, plotting_name)
plot_summary(this_parameter, yaxis_type = "log", text_size, 
             site_types, site_labels, site_colors)
plot_timeseries(this_parameter, yaxis_type = "log", text_size, 
                site_types, site_labels, site_colors)
```

## Silica, Dissolved (as SiO2)
[Back to top](#top)

Observations:

* Lake values may have a seasonal pattern
* Lake values are lower than groundwater values
* Downgradient values are lower than upgradinet values at Pleasant Lake, but all
groundwater is similar at Long and Plainfield.
* Levels are less variable in grundwater than in the lake

```{r SiO2, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=5, fig.align="center"}
parameter     <-"SILICA, DISSOLVED (MG/L AS SI02)"
plotting_name <- "Dissolved Silica as SiO2"

this_parameter <- filter_parameter(water_chem, parameter, plotting_name)
plot_summary(this_parameter, yaxis_type = "number", text_size, 
             site_types, site_labels, site_colors)
plot_timeseries(this_parameter, yaxis_type = "number", text_size, 
                site_types, site_labels, site_colors)
```

## pH
[Back to top](#top)

Observations:

* Lakes and groundwater are about as variable among sites as within sites
* Lake pH is higher than groundwater pH at Pleasant and Plainfield, but
lower at Long.
* There is an anomolously low value for LL-05 from the last sample date. 

```{r pH, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=5, fig.align="center"}
parameter     <-"PH LAB"
plotting_name <- "pH"

this_parameter <- filter_parameter(water_chem, parameter, plotting_name)
plot_summary(this_parameter, yaxis_type = "number", text_size, 
             site_types, site_labels, site_colors)
plot_timeseries(this_parameter, yaxis_type = "number", text_size, 
                site_types, site_labels, site_colors)
```


## Sulfate, Total
[Back to top](#top)

Observations:

* Sulfate is below the limit of quality for nearly all samples

```{r So4, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=5, fig.align="center"}
parameter     <- c("SULFATE TOTAL", "SULFATE DISS, AS SO4")
plotting_name <- "Sulfate"

this_parameter <- filter_parameter(water_chem, parameter, plotting_name)
plot_summary(this_parameter, yaxis_type = "number", text_size, 
             site_types, site_labels, site_colors)
plot_timeseries(this_parameter, yaxis_type = "number", text_size, 
                site_types, site_labels, site_colors)
```

