---
title: "CSLS Water Chem Visualization: Lakes Only"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{chem_lakes_only}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, echo=FALSE, warning=FALSE, message=FALSE}
library(knitr)
library(CSLSchem)
library(dplyr)
library(stringr)
library(lubridate)
library(reshape2)
library(extrafont)

text_size    <- 12
lakes        <- c("Pleasant", "Long", "Plainfield")
SWIMS        <- CSLSchem::SWIMS
SWIMS        <- SWIMS %>%
                filter(.data$site_type == "lake")
```

# Overview {#top}
The data used here includes data from both SWIMS and LDES. Duplicates between
the two data sources as well as samples labeled as blanks or duplicates are
removed.

## Calcium, Total Recoverable
[Back to top](#top)

```{r Ca, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=5, fig.align="center"}
parameter     <-"CALCIUM TOTAL RECOVERABLE"
plotting_name <- "Total Recoverable Calcium"

SWIMS_parameter <- filter_parameter(SWIMS, parameter, plotting_name)
plot_summary_by_year(SWIMS_parameter, text_size)
plot_timeseries_by_year(SWIMS_parameter, text_size)
```

## Magnesium, Total Recoverable
[Back to top](#top)

```{r Mg, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=5, fig.align="center"}
parameter     <- "MAGNESIUM TOTAL RECOVERABLE"
plotting_name <- "Total Recoverabel Magnesium"

SWIMS_parameter <- filter_parameter(SWIMS, parameter, plotting_name)
plot_summary_by_year(SWIMS_parameter, text_size)
plot_timeseries_by_year(SWIMS_parameter, text_size)
```

## Alkalinity, Total CaCO3
[Back to top](#top)

```{r CACO3, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=5, fig.align="center"}
parameter     <-"ALKALINITY TOTAL CACO3"
plotting_name <- "Alkalinity Total CaCO3"

SWIMS_parameter <- filter_parameter(SWIMS, parameter, plotting_name)
plot_summary_by_year(SWIMS_parameter, text_size)
plot_timeseries_by_year(SWIMS_parameter, text_size)
```


## Hardness, Total Recoverable Calculation
[Back to top](#top)

```{r hardness, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=5, fig.align="center"}
parameter     <-"HARDNESS TOTAL RECOVERABLE CALCULATION"
plotting_name <- "Total Recoverable Hardness"

SWIMS_parameter <- filter_parameter(SWIMS, parameter, plotting_name)
plot_summary_by_year(SWIMS_parameter, text_size)
plot_timeseries_by_year(SWIMS_parameter, text_size)
```

## Aluminum, Total Recoverable
[Back to top](#top)

```{r Al, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=5, fig.align="center"}
parameter     <-"ALUMINUM,TOTAL RECOVERABLE"
plotting_name <- "Total Recoverable Aluminum"

SWIMS_parameter <- filter_parameter(SWIMS, parameter, plotting_name)
plot_summary_by_year(SWIMS_parameter, text_size)
plot_timeseries_by_year(SWIMS_parameter, text_size)
```

## Iron, Total Recoverable
[Back to top](#top)


```{r Fe, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=5, fig.align="center"}
parameter     <-"IRON TOTAL RECOVERABLE"
plotting_name <- "Total Recoverable Iron"

SWIMS_parameter <- filter_parameter(SWIMS, parameter, plotting_name)
plot_summary_by_year(SWIMS_parameter, text_size)
plot_timeseries_by_year(SWIMS_parameter, text_size)
```

## Manganese, Total Recoverable
[Back to top](#top)

```{r Mn, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=5, fig.align="center"}
parameter     <-"MANGANESE, TOTAL RECOVERABLE"
plotting_name <- "Total Recoverable Manganese"

SWIMS_parameter <- filter_parameter(SWIMS, parameter, plotting_name)
plot_summary_by_year(SWIMS_parameter, text_size)
plot_timeseries_by_year(SWIMS_parameter, text_size)
```

## Conductivity
[Back to top](#top)

```{r Conductivity, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=5, fig.align="center"}
parameter     <-"CONDUCTIVITY, UMHOS/CM @ 25C"
plotting_name <- "Conductivity"

SWIMS_parameter <- filter_parameter(SWIMS, parameter, plotting_name)
plot_summary_by_year(SWIMS_parameter, text_size)
plot_timeseries_by_year(SWIMS_parameter, text_size)
```

## Chloride
[Back to top](#top)

```{r Cl, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=5, fig.align="center"}
parameter     <-"CHLORIDE"
plotting_name <- "Chloride"

SWIMS_parameter <- filter_parameter(SWIMS, parameter, plotting_name)
plot_summary_by_year(SWIMS_parameter, text_size)
plot_timeseries_by_year(SWIMS_parameter, text_size)
```

## Sodium, Total Recoverable
[Back to top](#top)

```{r Na, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=5, fig.align="center"}
parameter     <-"SODIUM TOTAL RECOVERABLE"
plotting_name <- "Total Recoverable Sodium"

SWIMS_parameter <- filter_parameter(SWIMS, parameter, plotting_name)
plot_summary_by_year(SWIMS_parameter, text_size)
plot_timeseries_by_year(SWIMS_parameter, text_size)
```

## Nitrogen, NH3-N Dissolved
[Back to top](#top)

```{r NH3, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=5, fig.align="center"}
parameter     <-"NITROGEN NH3-N DISS"
plotting_name <- "Dissolved Ammonium as Nitrogen"

SWIMS_parameter <- filter_parameter(SWIMS, parameter, plotting_name)
plot_summary_by_year(SWIMS_parameter, text_size)
plot_timeseries_by_year(SWIMS_parameter, text_size)
```

## Nitrogen, NO3 + NO2 Dissolved (as N)
[Back to top](#top)

```{r NO3NO2, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=5, fig.align="center"}
parameter     <-"NITROGEN NO3+NO2 DISS (AS N)"
plotting_name <- "Dissolved Nitrate/Nitrite as Nitrogen"

SWIMS_parameter <- filter_parameter(SWIMS, parameter, plotting_name)
plot_summary_by_year(SWIMS_parameter, text_size)
plot_timeseries_by_year(SWIMS_parameter, text_size)
```

## Phosphorus, Total
[Back to top](#top)

```{r P, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=5, fig.align="center"}
parameter     <-"PHOSPHORUS TOTAL"
plotting_name <- "Total Phosphorus"

SWIMS_parameter <- filter_parameter(SWIMS, parameter, plotting_name)
plot_summary_by_year(SWIMS_parameter, text_size)
plot_timeseries_by_year(SWIMS_parameter, text_size)
```

## Potassium, Total Recoverable
[Back to top](#top)

```{r K, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=5, fig.align="center"}
parameter     <-"POTASSIUM TOTAL RECOVERABLE"
plotting_name <- "Total Recoverable Potassium"

SWIMS_parameter <- filter_parameter(SWIMS, parameter, plotting_name)
plot_summary_by_year(SWIMS_parameter, text_size)
plot_timeseries_by_year(SWIMS_parameter, text_size)
```

## Silica, Dissolved (as SiO2)
[Back to top](#top)

```{r SiO2, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=5, fig.align="center"}
parameter     <-"SILICA, DISSOLVED (MG/L AS SI02)"
plotting_name <- "Dissolved Silica as SiO2"

SWIMS_parameter <- filter_parameter(SWIMS, parameter, plotting_name)
plot_summary_by_year(SWIMS_parameter, text_size)
plot_timeseries_by_year(SWIMS_parameter, text_size)
```

## pH
[Back to top](#top)

```{r pH, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=5, fig.align="center"}
parameter     <-"PH LAB"
plotting_name <- "pH"

SWIMS_parameter <- filter_parameter(SWIMS, parameter, plotting_name)
plot_summary_by_year(SWIMS_parameter, text_size)
plot_timeseries_by_year(SWIMS_parameter, text_size)
```


## Sulfate, Total
[Back to top](#top)

```{r So4, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=5, fig.align="center"}
parameter     <-"SULFATE TOTAL"
plotting_name <- "Total Sulfate"

SWIMS_parameter <- filter_parameter(SWIMS, parameter, plotting_name)
plot_summary_by_year(SWIMS_parameter, text_size)
plot_timeseries_by_year(SWIMS_parameter, text_size)
```

