---
title: "CSLS Water Chem Visualization"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Put the title of your vignette here}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, echo=FALSE, warning=FALSE, message=FALSE}
library(knitr)
library(CSLSchem)
library(dplyr)
library(stringr)
library(lubridate)
library(reshape2)

text_size    <- 12
lakes        <- c("Pleasant", "Long", "Plainfield")
SWIMS        <- classify_SWIMS_results(lakes)

find_and_plot <- function(parameter, plotting_name, text_size) {
  SWIMS_parameter <- filter_parameter(parameter, plotting_name)
  plot_obj        <- plot_overall(SWIMS_parameter, text_size)
  return(plot_obj)
}

plot_timeseries_fcn <- function(parameter, plotting_name, text_size, 
                                static_gw = TRUE) {
  SWIMS_parameter <- filter_parameter(parameter = parameter, 
                                      plotting_name = plotting_name,
                                      static_gw = static_gw)
  plot_obj <- plot_timeseries(SWIMS_parameter, text_size)
  return(plot_obj)
}

```

## Overview


## Alkalinity, Total CaCO3
```{r CACO3, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=5, fig.align="center"}
parameter     <-"ALKALINITY TOTAL CACO3"
plotting_name <- "Alkalinity Total CaCO3"

SWIMS_parameter <- filter_parameter(SWIMS, parameter, plotting_name)
plot_summary(SWIMS_parameter, text_size)
plot_timeseries(SWIMS_parameter, text_size)
```

## Aluminum, Total Recoverable
```{r Al, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=5, fig.align="center"}
parameter     <-"ALUMINUM,TOTAL RECOVERABLE"
plotting_name <- "Total Recoverable Aluminum"

SWIMS_parameter <- filter_parameter(SWIMS, parameter, plotting_name)
plot_summary(SWIMS_parameter, text_size)
plot_timeseries(SWIMS_parameter, text_size)
```

## Calcium, Total Recoverable
```{r Ca, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=5, fig.align="center"}
parameter     <-"CALCIUM TOTAL RECOVERABLE"
plotting_name <- "Total Recoverable Calcium"

SWIMS_parameter <- filter_parameter(SWIMS, parameter, plotting_name)
plot_summary(SWIMS_parameter, text_size)
plot_timeseries(SWIMS_parameter, text_size)
```

## Chloride

```{r Cl, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=5, fig.align="center"}
parameter     <-"CHLORIDE"
plotting_name <- "Chloride"

SWIMS_parameter <- filter_parameter(SWIMS, parameter, plotting_name)
plot_summary(SWIMS_parameter, text_size)
plot_timeseries(SWIMS_parameter, text_size)
```

## Conductivity

```{r Conductivity, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=5, fig.align="center"}
parameter     <-"CONDUCTIVITY, UMHOS/CM @ 25C"
plotting_name <- "Conductivity"

SWIMS_parameter <- filter_parameter(SWIMS, parameter, plotting_name)
plot_summary(SWIMS_parameter, text_size)
plot_timeseries(SWIMS_parameter, text_size)
```

## Hardness, Total Recoverable Calculation

```{r hardness, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=5, fig.align="center"}
parameter     <-"HARDNESS TOTAL RECOVERABLE CALCULATION"
plotting_name <- "Total Recoverable Hardness"

SWIMS_parameter <- filter_parameter(SWIMS, parameter, plotting_name)
plot_summary(SWIMS_parameter, text_size)
plot_timeseries(SWIMS_parameter, text_size)
```

## Iron, Total Recoverable

```{r Fe, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=5, fig.align="center"}
parameter     <-"IRON TOTAL RECOVERABLE"
plotting_name <- "Total Recoverable Iron"

SWIMS_parameter <- filter_parameter(SWIMS, parameter, plotting_name)
plot_summary(SWIMS_parameter, text_size)
plot_timeseries(SWIMS_parameter, text_size)
```

## Magnesium, Total Recoverable

```{r Mg, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=5, fig.align="center"}
parameter     <- "MAGNESIUM TOTAL RECOVERABLE"
plotting_name <- "Total Recoverabel Magnesium"

SWIMS_parameter <- filter_parameter(SWIMS, parameter, plotting_name)
plot_summary(SWIMS_parameter, text_size)
plot_timeseries(SWIMS_parameter, text_size)
```

## Manganese, Total Recoverable

```{r Mn, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=5, fig.align="center"}
parameter     <-"MANGANESE, TOTAL RECOVERABLE"
plotting_name <- "Total Recoverable Manganese"

SWIMS_parameter <- filter_parameter(SWIMS, parameter, plotting_name)
plot_summary(SWIMS_parameter, text_size)
plot_timeseries(SWIMS_parameter, text_size)
```

## Nitrogen, NH3-N Dissolved

```{r NH3, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=5, fig.align="center"}
parameter     <-"NITROGEN NH3-N DISS"
plotting_name <- "Dissolved Ammonium as Nitrogen"

SWIMS_parameter <- filter_parameter(SWIMS, parameter, plotting_name)
plot_summary(SWIMS_parameter, text_size)
plot_timeseries(SWIMS_parameter, text_size)
```

## Nitrogen, NO3 + NO2 Dissolved (as N)

```{r NO3NO2, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=5, fig.align="center"}
parameter     <-"NITROGEN NO3+NO2 DISS (AS N)"
plotting_name <- "Dissolved Nitrate/Nitrite as Nitrogen"

SWIMS_parameter <- filter_parameter(SWIMS, parameter, plotting_name)
plot_summary(SWIMS_parameter, text_size)
plot_timeseries(SWIMS_parameter, text_size)
```

## pH

```{r pH, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=5, fig.align="center"}
parameter     <-"PH LAB"
plotting_name <- "pH"

SWIMS_parameter <- filter_parameter(SWIMS, parameter, plotting_name)
plot_summary(SWIMS_parameter, text_size)
plot_timeseries(SWIMS_parameter, text_size)
```

## Phosphorus, Total

```{r P, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=5, fig.align="center"}
parameter     <-"PHOSPHORUS TOTAL"
plotting_name <- "Total Phosphorus"

SWIMS_parameter <- filter_parameter(SWIMS, parameter, plotting_name)
plot_summary(SWIMS_parameter, text_size)
plot_timeseries(SWIMS_parameter, text_size)
```

## Potassium, Total Recoverable

```{r K, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=5, fig.align="center"}
parameter     <-"POTASSIUM TOTAL RECOVERABLE"
plotting_name <- "Total Recoverable Potassium"

SWIMS_parameter <- filter_parameter(SWIMS, parameter, plotting_name)
plot_summary(SWIMS_parameter, text_size)
plot_timeseries(SWIMS_parameter, text_size)
```

## Silica, Dissolved (as SiO2)

```{r SiO2, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=5, fig.align="center"}
parameter     <-"SILICA, DISSOLVED (MG/L AS SI02)"
plotting_name <- "Dissolved Silica as SiO2"

SWIMS_parameter <- filter_parameter(SWIMS, parameter, plotting_name)
plot_summary(SWIMS_parameter, text_size)
plot_timeseries(SWIMS_parameter, text_size)
```

## Sodium, Total Recoverable

```{r Na, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=5, fig.align="center"}
parameter     <-"SODIUM TOTAL RECOVERABLE"
plotting_name <- "Total Recoverable Sodium"

SWIMS_parameter <- filter_parameter(SWIMS, parameter, plotting_name)
plot_summary(SWIMS_parameter, text_size)
plot_timeseries(SWIMS_parameter, text_size)
```

## Sulfate, Total

```{r So4, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=5, fig.align="center"}
parameter     <-"SULFATE TOTAL"
plotting_name <- "Total Sulfate"

SWIMS_parameter <- filter_parameter(SWIMS, parameter, plotting_name)
plot_summary(SWIMS_parameter, text_size)
plot_timeseries(SWIMS_parameter, text_size)
```

